Sentence Generation as Planning with Probabilistic LTAG\\

### 2010-05-20 trunk/Papers/pcrisp-10/introduction.tex
Related Work

While CRISP is efficient enough to perform well on large-scale
grammars \cite{kollerpetrick2009}, such grammars tend to offer many
different ways to express them same semantic representation.  This
makes it necessary for the generation system to be able to compute not
just grammatical sentences, but to identify which of these sentences
are the \emph{best}. This problem is exacerbated when using
treebank-derived grammars, which tend to underspecify the actual
constraints on grammaticality and instead rely on statistical
information learned from the treebank.  Indeed, there have been a
number of systems for statistical generation, which can exploit such
information to rank sentences appropriately
\cite{langkildeknight1998,whitebaldridge2003,belz2008}.  However,
these systems are restricted to performing surface realization.

In this paper, we bring these two strands of research together for the
first time.  We present the PCRISP system, which redefines the SPUD
generation problem in terms of probabilistic TAG grammars (PTAG,
\cite{resnik1992}) and then extends CRISP to solving the probabilistic
SPUD generation problem using metric planning \cite{hoffmann2003}.  We
evaluate PCRISP on a PTAG treebank extracted from the Wall Street
Journal Corpus \cite{chenschanker2004}.  While the evaluation results
are not competitive in terms of efficiency and suffer from data
sparseness problems, we still believe our system is a first step
towards probabilistic integrated generation.

>>>>

While CRISP is efficient enough to perform well on large-scale
grammars \cite{kollerpetrick2009}, such grammars tend to offer many
different ways to express the same semantic representation.  This
makes it necessary for the generation system to be able to compute not
just grammatical sentences, but to identify which of these sentences
are the \emph{best}. This problem is exacerbated when using
treebank-derived grammars, which tend to underspecify the actual
constraints on grammaticality and instead rely on statistical
information learned from the treebank.  Indeed, there have been a
number of systems for statistical generation, which can exploit such
information to rank sentences appropriately
\cite{langkildeknight1998,whitebaldridge2003,belz2008}.  However,
these systems are restricted to performing surface realization.

In this paper, we bring these two strands of research together for the
first time.  We present the PCRISP system, which redefines the SPUD
generation problem in terms of probabilistic TAG grammars (PTAG,
\cite{resnik1992}) and then extends CRISP to solving the probabilistic
SPUD generation problem using metric planning \cite{fox2002,hoffmann2003}.  We
evaluate PCRISP on a PTAG treebank extracted from the Wall Street
Journal Corpus \cite{chenschanker2004}.  While the evaluation results
are not competitive in terms of efficiency and suffer from data
sparseness problems, we still believe our system is a first step
towards probabilistic integrated generation.

### 2010-05-20 trunk/Papers/pcrisp-10/related.tex
Typos.
Related Work

\section{Related Work}
\label{sec:related}

Previous statistical NLG systems (e.g. \newcite{knight1995}, {\sc nitrogen} \cite{langkildeknight1998}, {\sc openCCG} \cite{whitebaldridge2003}
and the {\sc ltag} based realizer {\sc fergus} \cite{bangalorerambow2000}), follow a generate-and-select strategy using n-gram models over output structures
for discrimination. But since this approach focuses on discriminating realization output for the same sentence plan only, it is incompatible with the
integrated treatment of sentence planning and realization.

One of the few systems uses generative models for sentence generation is {\sc pCru} \newcite{belz2008}. However, her model is defined on possible states of the generation system itself, not on the underlying grammar formalism.

$\cdots$

>>>>

\section{Related Work}
\label{sec:related}
Statistical methods are extremely popular for surface realization, but have not been used in systems that integrate sentence planning. Most statistical generation approaches follow a generate-and-select strategy, first proposed by \newcite{knight1995} in their {\sc nitrogen} system. Such systems generate a set of candidate sentences using a conventional (non-statistical) grammar and then select the best output sentence by applying a statistical language model. This family includes systems such as {\sc HAL}ogen \cite{langkildeknight1998,langkilde2000} and open{\sc CCG} \cite{whitebaldridge2003}.  The {\sc fergus} system by \newcite{bangalorerambow2000} is a variant of this approach that, like {\sc pcrisp}, employs {\sc ltag}. It first assigns elementary trees to each entity in the input sentence plan using a statistical tree model and then computes the most likely derivation using using only these trees with an $n$-gram model on the output sentence. An alternative to the generate and select approach is to use a probabilistic grammar model, like {\sc ptag}, trained on automatic parses \cite{zhongstent2005}. A related approach uses a model over local decisions of the generation system itself \cite{belz2008}. Both model can either be used to discriminate a set of output candidates, or more directly to choose the next best decision locally. Our approach is similar, because it uses {\sc ptag} to find the most likely output structure. However, the previous work discussed so far addresses surface realization only.
The generate and select approach is even incompatible with the integrated treatment of sentence planning and realization, because it discriminates between possible linearization of the same lexical material. In contrast, {\sc pcrisp} addresses sentence planning as well.

Our treatment of integrated sentence planning and surface realization as planning, is inherited directly from {\sc crisp} \cite{kollerstone2007}.  Planning has long played a role in generation, but has focused on discourse planning instead of specifically addressing sentence generation \cite{hovy1988,appelt1992}. The applicability of these ideas was limited at that time because efficient planning technology was not available. Recently the development of more efficient planning algorithms (e.g. \newcite{hoffmannnebel2001,richterwestphal2008}) spawned a renewed interest in planning for NLG.  
{\sc Crisp} uses such efficient algorithms to efficiently solve the sentence generation problem defined by {\sc spud} \cite{stonedoran1997}. {\sc Spud}, which instead used an incomplete greedy algorithm, is based on an {\sc ltag} whose trees are augmented with semantic and pragmatic constraints. Given a communicative goal, a solution to the {\sc spud} problem realizes this goal and simultaneosly selects refering expressions. The next section explores this idea in more detail.
